\begin{topic}{abelian-category}{abelian category}
    An \textbf{abelian category} is a category such that
    \begin{itemize}
        \item every $\Hom(A, B)$ is an abelian group, and composition is bilinear,
        \item it has a zero object,
        \item it has all binary biproducts,
        \item it has all kernels and cokernels,
        \item every monomoprhism is the kernel of some morphism, and every epimoprhism is the cokernel of some morphism.
    \end{itemize}
\end{topic}

\begin{topic}{five-lemma}{five lemma}
    Let $\mathcal{A}$ be an abelian category, and consider the following commutative diagram.
    \[ \begin{tikzcd} A \arrow{r}{f} \arrow{d}{a} & B \arrow{r}{g} \arrow{d}{b} & C \arrow{r}{h} \arrow{d}{c} & D \arrow{r}{i} \arrow{d}{d} & E \arrow{d}{e} \\ A' \arrow{r}{f'} & B' \arrow{r}{g'} & C' \arrow{r}{h'} & D \arrow{r}{i'} & E' \end{tikzcd} \]
    Suppose that both rows are exact. The \textbf{five lemma} states that
    \begin{itemize}
        \item if $b$ and $d$ are injective and $a$ is surjective, then $c$ is injective.
        
        \item if $b$ and $d$ are surjective and $e$ is injective, then $c$ is surjective.
        
        \item if $b$ and $d$ are isomorphisms, $a$ is surjective and $e$ is injective, then $c$ is an isomorphism.
    \end{itemize}
\end{topic}

\begin{topic}{snake-lemma}{snake lemma}
    Let $\mathcal{A}$ be an abelian category, and consider the following commutative diagram.
    \[ \begin{tikzcd} & A \arrow{d}{a} \arrow{r}{f} & B \arrow{d}{b} \arrow{r}{g} & C \arrow{d}{c} \arrow{r} & 0 \\ 0 \arrow{r} & A' \arrow{r}{f'} & B' \arrow{r}{g'} & C' \end{tikzcd} \]
    Suppose that the rows are exact. The \textbf{snake lemma} states that there is an exact sequence of kernels and cokernels
    \[ \ker a \rightarrow \ker b \rightarrow \ker c \xrightarrow{d} \coker a \rightarrow \coker b \rightarrow \coker c \]
    where $d$ is known as the \textit{connecting homomorphism}. Furthermore, if $f$ is \tref{CT:monomorphism}{mono} then so is $\ker a \to \ker b$, and if $g'$ is \tref{CT:epimorphism}{epi} then so is $\coker b \to \coker c$.
\end{topic}

\begin{topic}{dg-category}{dg-category}
    A \textbf{differential graded category} is a category whose morphism sets are \tref{AT:cochain-complex}{cochain complexes}, that is
    \[ \Hom(A, B) = \bigoplus_{n \in \ZZ} \Hom_n(A, B) \]
    with differential $d : \Hom_n(A, B) \to \Hom_{n + 1}(A, B)$ such that $d^2 = 0$. Furthermore, composition of morphisms
    \[ \Hom(A, B) \otimes \Hom(B, C) \to \Hom(A, C) \]
    must be a \tref{AT:chain-map}{chain map} and $d(\id_A) = 0$.
\end{topic}
